04-Data-Analysis

### 2021-04-22 00-Data-Processing/04-Data-Analysis.Rmd
Analysis of accessibility by age is probably complete

Load packages:
```{r load-packages, include=FALSE}
library(patchwork)
library(sf)
library(tidyverse)
#library(vaccHamilton)
```

@@

ggplot(accessibility_test, 
       aes(x = threshold, y = accessibility)) +
  geom_line()
```

@@

Join geometry:
```{r, add-geometry-to-results, include=FALSE}
los <- los %>%
  left_join(foodbank_locations %>%
              dplyr::select(ID),
            by = c("OBJECTID" = "ID")) %>%
  st_as_sf()

accessibility <- accessibility %>%
  left_join(dwelling_network_points_2016 %>%
              dplyr::select(ID),
            by = c("UID" = "ID")) %>%
  st_as_sf()
```

@@

Plot accessibility before and during covid:
```{r}
acc_pre <- ggplot() +
  geom_sf(data = hamilton_boundary) +
  geom_sf(data = accessibility %>%
            filter(timing == "Pre-covid"),
          aes(color = accessibility)) +
  scale_color_distiller(palette = "OrRd", 
                        direction = 1)

acc_covid <- ggplot() +
  geom_sf(data = hamilton_boundary) +
  geom_sf(data = accessibility %>%
            filter(timing == "During covid"),
          aes(color = accessibility)) +
  scale_color_distiller(palette = "OrRd", 
                        direction = 1)

@@

Plot the change in accessibility by DA:
```{r}
accessibility %>%
  st_drop_geometry() %>%
  # Copy the DA identifier to the accessibility table
  left_join(dwelling_network_points_2016 %>%
              st_drop_geometry() %>%
              transmute(UID = ID, DAUID),
            by = "UID") %>%
  # Group by DA and time
  group_by(DAUID, timing) %>%
  # Calculate the accessibility by DA before and during covid
  summarize(accessibility = sum(accessibility),
            timing = first(timing),
            .groups = "drop") %>%
  # Pivot wider to calculate the differences in accessibility from pre-covid to covid
  pivot_wider(names_from = timing, 
              values_from = accessibility) %>%
  # Calculate changes in accessibility
  mutate(acc_d = `During covid` / `Pre-covid` - 1) %>%
  # Join DA geometry
  left_join(data_da_2016 %>%
            transmute(DAUID = GeoUID,
                      geometry),
          by = "DAUID") %>%
  # Convert to simple features
  st_as_sf() %>%
  # Plot
  ggplot() +
  geom_sf(data = hamilton_boundary) +
  geom_sf(aes(fill = acc_d),
          color = NA) +
  scale_fill_distiller(palette = "OrRd", 
                        direction = -1)

@@

@@

Get minimum travel times in the pre-covid scenario:

```{r minimum-travel-time-baseline, include=FALSE}
# retrieve minimum travel time for pre-covid locations
min_ttm_car_baseline <- ttm_car %>%
  #dplyr::filter(COVIDClose == "No") %>%
  filter(TYPE != "Food Banks") %>%
  group_by(UID) %>%
  summarize(travel_time = min(travel_time),
            DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            p_car = first(p_car),
            p_transit = first(p_transit),
            p_walk = first(p_walk),
            children = first(children),
            adults = first(adults),
            seniors = first(seniors),
            .groups = "drop") %>%

@@

Bind the tables and label by mode:

```{r weighted-travel-time-by-mode-baseline, include=FALSE}
# Baseline analysis (pre-COVID)
## Bind the three tables and label them by mode:
min_ttm_mode_baseline <- rbind(data.frame(min_ttm_car_baseline, Mode = "Car"),
                               data.frame(min_ttm_transit_baseline, Mode = "Transit"),
                               data.frame(min_ttm_walk_baseline, Mode = "Walking"))
```

```{r aggregate-weighted-travel-times-baseline, include=FALSE}
# Baseline analysis (pre-COVID)
## Aggregate weighted travel times by mode:
min_ttm_mode_baseline <- min_ttm_mode_baseline %>%
  group_by(UID) %>%
  summarize(DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            children = first(children),
            adults = first(adults),
            seniors = first(seniors),
            weighted_travel_time = sum(weighted_travel_time)) %>%

@@

```{r summarize-results-by-taz-baseline, include=FALSE}
# Baseline analysis (pre-COVID)
## Summarize by TAZ and join geometry:
min_ttm_baseline_taz <- min_ttm_mode_baseline %>%
  group_by(TAZUID) %>%
  summarize(weighted_travel_time = mean(weighted_travel_time),
            wph_children = sum(wph_children),
            wph_adults = sum(wph_adults),
            wph_seniors = sum(wph_seniors)) %>%

@@

```{r}
tm_shape(min_ttm_baseline_taz) +
  tm_fill("wph_children")

@@

```{r weighted-travel-time-map-baseline}
# Baseline analysis (pre-COVID)
## Map weighted travel time:
ttm_map_baseline <- ggplot() +
  geom_sf(data = min_ttm_baseline_taz,
          aes(fill = weighted_travel_time),
          color = NA) +
  scale_fill_distiller(name = "Expected travel time (min)",
                       palette = "OrRd", 
                       direction = 1)

@@

```{r minimum-travel-time-covid, include=FALSE}
# retrieve minimum travel time during covid to sites that didnt close
min_ttm_car_covid <- ttm_car %>%
  dplyr::filter(COVIDClose == "No", TYPE != "Food Banks") %>%
  group_by(UID) %>%
  summarize(travel_time = min(travel_time),
            DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            p_car = first(p_car),
            p_transit = first(p_transit),
            p_walk = first(p_walk),
            children = first(children),
            adults = first(adults),
            seniors = first(seniors),
            .groups = "drop") %>%

@@

Bind the tables and label by mode:

```{r weighted-travel-time-by-mode-covid, include=FALSE}
# Baseline analysis (pre-COVID)
## Bind the three tables and label them by mode:
min_ttm_mode_covid <- rbind(data.frame(min_ttm_car_covid, Mode = "Car"),
                               data.frame(min_ttm_transit_covid, Mode = "Transit"),
                               data.frame(min_ttm_walk_covid, Mode = "Walking"))
```

```{r aggregate-weighted-travel-times-covid, include=FALSE}
# Baseline analysis (pre-COVID)
## Aggregate weighted travel times by mode:
min_ttm_mode_covid <- min_ttm_mode_covid %>%
  group_by(UID) %>%
  summarize(DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            children = first(children),
            adults = first(adults),
            seniors = first(seniors),
            weighted_travel_time = sum(weighted_travel_time)) %>%

@@

```{r summarize-results-by-taz-covid, include=FALSE}
# Baseline analysis (PILOT LOCATIONS ONLY)
## Summarize by TAZ and join geometry:
min_ttm_covid_taz <- min_ttm_mode_covid %>%
  group_by(TAZUID) %>%
  summarize(weighted_travel_time = mean(weighted_travel_time),
            wph_children = sum(wph_children),
            wph_adults = sum(wph_adults),
            wph_seniors = sum(wph_seniors)) %>%

@@

@@

```{r weighted-travel-time-map-covid}
# Baseline analysis (PILOT LOCATIONS ONLY)
## Map weighted travel time:
ttm_map_covid <- ggplot() +
  geom_sf(data = min_ttm_covid_taz,
          aes(fill = weighted_travel_time),
          color = NA) +
  scale_fill_distiller(name = "Expected travel time (min)",
                       palette = "OrRd", 
                       direction = 1)

@@

Join the min travel time tables together

```{r}
min_ttm_d <- min_ttm_baseline_taz %>% 
  left_join(min_ttm_covid_taz %>%
              st_drop_geometry() %>%
              transmute(TAZUID,
                        wph_children_covid = wph_children,
                        wph_adults_covid = wph_adults,
                        wph_seniors_covid = wph_seniors), 
            by = "TAZUID") %>%
  mutate(wph_children_d = wph_children_covid / wph_children - 1,
         wph_children_d = replace_na(wph_children_d, 0),
         wph_adults_d = wph_adults_covid / wph_adults - 1,
         wph_adults_d = replace_na(wph_adults_d, 0),
         wph_seniors_d = wph_seniors_covid / wph_seniors - 1,
         wph_seniors_d = replace_na(wph_seniors_d, 0))

>>>>

Load packages:
```{r load-packages, include=FALSE}
library(patchwork)
library(sf)
library(tidyverse)
library(tmap)
#library(vaccHamilton)
```

@@

acc_sensitivity <- ggplot(accessibility_test, 
       aes(x = threshold, y = accessibility)) +
  geom_line() +
  ylab("Accessibility")

acc_sensitivity
```

@@

Join geometry:
```{r, add-geometry-to-level-of-service, include=FALSE}
los <- los %>%
  left_join(foodbank_locations %>%
              dplyr::select(ID),
            by = c("OBJECTID" = "ID")) %>%
  st_as_sf()
```

@@

Join zonal identifiers to accessibility results:
```{r}
accessibility <- accessibility %>%
  left_join(dwelling_network_points_2016 %>%
            st_drop_geometry() %>%
              dplyr::select(ID,
                            DAUID,
                            TAZUID,
                            children,
                            adults,
                            seniors),
            by = c("UID" = "ID")) %>%
  left_join(population_age_income %>%
              st_drop_geometry() %>%
              transmute(TAZUID),
            by = "TAZUID") 
```

Accessibility by TAZ:
```{r}  
accessibility_taz <- accessibility %>%
  group_by(TAZUID, 
           timing) %>%
  summarize(accessibility = sum(accessibility),
            timing = first(timing),
            .groups = "drop") %>%
  left_join(population_age_income %>%
              transmute(TAZUID,
                        children,
                        adults,
                        seniors),
            by = "TAZUID") %>%
  st_as_sf()
```

Plot accessibility before and during covid:
```{r}
acc_pre <- ggplot() +
  geom_sf(data = hamilton_boundary) +
  geom_sf(data = accessibility_taz %>%
            filter(timing == "Pre-covid"),
          aes(fill = accessibility)) +
  scale_fill_distiller(palette = "OrRd", 
                        direction = 1)

acc_covid <- ggplot() +
  geom_sf(data = hamilton_boundary) +
  geom_sf(data = accessibility_taz %>%
            filter(timing == "During covid"),
          aes(fill = accessibility)) +
  scale_fill_distiller(palette = "OrRd", 
                        direction = 1)

@@

Calculate accessibility change by TAZ:
```{r}
accessibilty_taz_change <- accessibility_taz %>%
  st_drop_geometry() %>%
  pivot_wider(names_from = timing, 
              values_from = accessibility) %>%
  # Calculate changes in accessibility
  #mutate(acc_d = `During covid` / `Pre-covid` - 1) %>% # Percentage change is rather flat and not as easy to interpret as the difference
  mutate(acc_d = `During covid` - `Pre-covid`) %>%
  # Join TAZ geometry
  left_join(population_age_income %>%
            transmute(TAZUID,
                      geometry),
          by = "TAZUID") %>%
  # Convert to simple features
  st_as_sf()
```

Plot the change in accessibility by TAZ:
```{r}
  # Plot
  ggplot() +
  geom_sf(data = accessibilty_taz_change,
          aes(fill = acc_d),
          color = NA) +
  geom_sf(data = population_age_income,
          fill = NA) +
  scale_fill_fermenter(palette = "OrRd", 
                        direction = -1)
```

@@

Calculate accessibility change by parcel:
```{r}
accessibility_change <- accessibility %>%
  select(UID, timing, accessibility) %>%
  pivot_wider(names_from = timing, 
              values_from = accessibility,
              values_fill = 0) %>%
  # Calculate changes in accessibility
  #mutate(acc_d = `During covid` / `Pre-covid` - 1) %>% # Percentage change is rather flat and not as easy to interpret as the difference
  mutate(acc_d = `During covid` - `Pre-covid`)
```

Join accessibility change to accessibility table:
```{r}
accessibility <- accessibility %>%
  left_join(accessibility_change %>%
              select(UID, acc_d),
            by = "UID")
```

Quartiles of accessibility:
```{r}
# acc_quintiles_pre <- quantile(accessibility %>%
#                                 filter(timing == "Pre-covid") %>%
#                                 pull(accessibility),
#                               c(0, 0.25, 0.5, 0.75, 1))
# acc_quintiles_covid <- quantile(accessibility %>%
#                                 filter(timing == "During covid") %>%
#                                 pull(accessibility),
#                               c(0, 0.25, 0.5, 0.75, 1))
```

Quartiles of accessibility change:
```{r}
acc_quintiles_change <- quantile(accessibilty_change$acc_d,
                              c(0, 0.25, 0.5, 0.75, 1))
```

Add labels of accessibility levels to table:
```{r}
# accessibility <- accessibility %>%
#   mutate(accessibility_level = case_when(timing == "Pre-covid" & accessibility <= acc_quintiles_pre[2] ~ "Bottom 25%",
#                                          timing == "Pre-covid" & accessibility > acc_quintiles_pre[2] & accessibility <= acc_quintiles_pre[3] ~ "Third 25%",
#                                          timing == "Pre-covid" & accessibility > acc_quintiles_pre[3] & accessibility <= acc_quintiles_pre[4] ~ "Second 25%",
#                                          timing == "Pre-covid" & accessibility > acc_quintiles_pre[4] ~ "Top 25%",
#                                          timing == "During covid" & accessibility <= acc_quintiles_covid[2] ~ "Bottom 25%",
#                                          timing == "During covid" & accessibility > acc_quintiles_covid[2] & accessibility <= acc_quintiles_covid[3] ~ "Third 25%",
#                                          timing == "During covid" & accessibility > acc_quintiles_covid[3] & accessibility <= acc_quintiles_covid[4] ~ "Second 25%",
#                                          timing == "During covid" & accessibility > acc_quintiles_covid[4] ~ "Top 25%"),
#            accessibility_level = factor(accessibility_level,
#                                         levels = c("Top 25%", "Second 25%", "Third 25%", "Bottom 25%"),
#                                         ordered = TRUE))
```

Add labels of accessibility loss to table:
```{r}
accessibility <- accessibility %>%
  mutate(accessibility_level = case_when(acc_d <= acc_quintiles_change[2] ~ "Bottom 25%",
                                         acc_d > acc_quintiles_change[2] & acc_d <= acc_quintiles_change[3] ~ "Third 25%",
                                         acc_d > acc_quintiles_change[3] & acc_d <= acc_quintiles_change[4] ~ "Second 25%",
                                         acc_d > acc_quintiles_change[4] ~ "Top 25%",
                                         accessibility <= acc_quintiles_covid[2] ~ "Bottom 25%"),
           accessibility_level = factor(accessibility_level,
                                        levels = c("Top 25%", "Second 25%", "Third 25%", "Bottom 25%"),
                                        ordered = TRUE))
```

Calculate accessibility by age group and accessibility level:
```{r}
accessibility_by_age <- accessibility %>%
  group_by(timing, 
           accessibility_level) %>%
  summarize(children = sum(children),
            adults = sum(adults),
            seniors = sum(seniors),
            accessibility = sum(accessibility),
            .groups = "drop")

accessibility_by_age
```

The accessibility level is the magnitude of the accessibility loss: the top 25% faced the smallest losses in accessibility, and the bottom 25% faced the greatest losses in accessibility.

```{r}
cbind(accessibility_by_age %>% 
        filter(timing == "Pre-covid") %>% 
        select(children:accessibility) %>% 
        rename(accessibility_pre = accessibility),
      accessibility_by_age %>% 
        filter(timing == "During covid") %>% 
        transmute(accessibility_covid = accessibility)) %>%
  mutate(change = accessibility_covid - accessibility_pre)
```

Sanity check:
```{r}
accessibility_by_age %>%
  group_by(timing) %>%
  summarize(children = sum(children),
            adults = sum(adults),
            seniors = sum(seniors),
            accessibility = sum(accessibility))
```

@@

Get weighted minimum travel times in the pre-covid scenario:
```{r minimum-travel-time-baseline, include=FALSE}
# retrieve minimum travel time for pre-covid locations
min_ttm_car_baseline <- ttm_car %>%
  #dplyr::filter(COVIDClose == "No") %>%
  filter(TYPE != "Food Banks") %>%
  group_by(UID) %>%
  summarize(travel_time = min(travel_time),
            DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            p_car = first(p_car),
            p_transit = first(p_transit),
            p_walk = first(p_walk),
            children = first(children),
            adults = first(adults),
            seniors = first(seniors),
            .groups = "drop") %>%

@@

Bind the tables and label by mode:
```{r weighted-travel-time-by-mode-baseline, include=FALSE}
# Baseline analysis (pre-COVID)
## Bind the three tables and label them by mode:
min_ttm_mode_baseline <- rbind(data.frame(min_ttm_car_baseline, Mode = "Car"),
                               data.frame(min_ttm_transit_baseline, Mode = "Transit"),
                               data.frame(min_ttm_walk_baseline, Mode = "Walking"))
```

Calculate the weighted person hours of travel by age group:
```{r aggregate-weighted-travel-times-baseline, include=FALSE}
# Baseline analysis (pre-COVID)
## Aggregate weighted travel times by mode:
min_ttm_mode_baseline <- min_ttm_mode_baseline %>%
  group_by(UID) %>%
  summarize(DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            children = first(children),
            adults = first(adults),
            seniors = first(seniors),
            # Weighted travel time is the sum of the travel times weighted by the proportion of trips by mode
            weighted_travel_time = sum(weighted_travel_time)) %>%

@@

Calculate the weighted person hours of travel by TAZ by population group:
```{r summarize-results-by-taz-baseline, include=FALSE}
# Baseline analysis (pre-COVID)
## Summarize by TAZ and join geometry:
min_ttm_baseline_taz <- min_ttm_mode_baseline %>%
  group_by(TAZUID) %>%
  summarize(weighted_travel_time = mean(weighted_travel_time),
            wph_children = sum(wph_children),
            wph_adults = sum(wph_adults),
            wph_seniors = sum(wph_seniors)) %>%

@@

Plot weighted travel time to nearest facility:
```{r weighted-travel-time-map-baseline}
# Baseline analysis (pre-COVID)
## Map weighted travel time:
ttm_map_baseline <- ggplot() +
  geom_sf(data = min_ttm_baseline_taz,
          aes(fill = weighted_travel_time),
          color = NA) +
  scale_fill_fermenter(name = "Expected travel time (min)",
                       palette = "OrRd", 
                       direction = 1)

Plot the weighted person hours of travel for the population groups in low income households (travel times in minutes):
```{r}
tm_shape(min_ttm_baseline_taz) +
  tm_fill("wph_children")

@@

@@

Calculate weighted travel time to the nearest facilty by mode:
```{r minimum-travel-time-covid, include=FALSE}
# retrieve minimum travel time during covid to sites that didn't close
min_ttm_car_covid <- ttm_car %>%
  dplyr::filter(COVIDClose == "No", TYPE != "Food Banks") %>%
  group_by(UID) %>%
  summarize(travel_time = min(travel_time),
            DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            p_car = first(p_car),
            p_transit = first(p_transit),
            p_walk = first(p_walk),
            children = first(children),
            adults = first(adults),
            seniors = first(seniors),
            .groups = "drop") %>%

@@

Bind the tables and label by mode:
```{r weighted-travel-time-by-mode-covid, include=FALSE}
# Baseline analysis (pre-COVID)
## Bind the three tables and label them by mode:
min_ttm_mode_covid <- rbind(data.frame(min_ttm_car_covid, Mode = "Car"),
                               data.frame(min_ttm_transit_covid, Mode = "Transit"),
                               data.frame(min_ttm_walk_covid, Mode = "Walking"))
```

Calculate the weighted person hours of travel by age group:
```{r aggregate-weighted-travel-times-covid, include=FALSE}
# Baseline analysis (pre-COVID)
## Aggregate weighted travel times by mode:
min_ttm_mode_covid <- min_ttm_mode_covid %>%
  group_by(UID) %>%
  summarize(DAUID = first(DAUID),
            TAZUID = first(TAZUID),
            children = first(children),
            adults = first(adults),
            seniors = first(seniors),
            weighted_travel_time = sum(weighted_travel_time)) %>%

@@

Aggregate the number of weighted person hours of travel by age for each TAZ:
```{r summarize-results-by-taz-covid, include=FALSE}
# Baseline analysis (PILOT LOCATIONS ONLY)
## Summarize by TAZ and join geometry:
min_ttm_covid_taz <- min_ttm_mode_covid %>%
  group_by(TAZUID) %>%
  summarize(weighted_travel_time = mean(weighted_travel_time),
            wph_children = sum(wph_children),
            wph_adults = sum(wph_adults),
            wph_seniors = sum(wph_seniors)) %>%

@@

Map weighted travel time to nearest facility:
```{r weighted-travel-time-map-covid}
# Baseline analysis (PILOT LOCATIONS ONLY)
## Map weighted travel time:
ttm_map_covid <- ggplot() +
  geom_sf(data = min_ttm_covid_taz,
          aes(fill = weighted_travel_time),
          color = NA) +
  scale_fill_distiller(name = "Expected travel time (min)",
                       palette = "OrRd", 
                       direction = 1)

@@

@@

Join the min travel time tables together:
```{r}
min_ttm_d <- min_ttm_baseline_taz %>% 
  left_join(min_ttm_covid_taz %>%
              st_drop_geometry() %>%
              transmute(TAZUID,
                        wph_children_covid = wph_children,
                        wph_adults_covid = wph_adults,
                        wph_seniors_covid = wph_seniors), 
            by = "TAZUID") %>%
  mutate(wph_children_d = wph_children_covid / wph_children - 1,
         wph_children_d = replace_na(wph_children_d, 0),
         wph_adults_d = wph_adults_covid / wph_adults - 1,
         wph_adults_d = replace_na(wph_adults_d, 0),
         wph_seniors_d = wph_seniors_covid / wph_seniors - 1,
         wph_seniors_d = replace_na(wph_seniors_d, 0))
